---
title: 'Reproducible Research: Peer Assessment 1'
author: 'Cole Yorston'
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
 ```
 

## Loading and preprocessing the data

 ```{r}
library(knitr)
library(ggplot2)
library(tinytex)

if (!file.exists('activity.csv')) {
  
  unzip('./activity.zip', exdir = '.');
}
activity <- read.csv('activity.csv');
summary(activity)
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
activity_no_nulls <- activity[with(activity, { !(is.na(steps)) } ), ]
steps_by_day <- aggregate(steps ~ date, activity_no_nulls, sum)
hist(steps_by_day$steps, col="grey", xlab = "Number of Steps Taken", main= "Total Steps Taken by Day")

```

# Imported libraries and assessed NA values, file is unzipped

## What is mean total number of steps taken per day?

```{r}
mean(steps_by_day$steps)
median(steps_by_day$steps)

# Mean value is determined here. 
```

## What is the average daily activity pattern?
```{r}
avg_steps_interval <- aggregate(steps ~ interval, activity, mean)
plot(avg_steps_interval$interval, avg_steps_interval$steps, type='l', col=1, main="Average Steps per Day by Interval", xlab="Time Intervals", ylab="Average number of steps")
interval_max <- which.max(avg_steps_interval$steps)
interval_max_ <- avg_steps_interval[interval_max, ]$interval
interval_max__ <- round(avg_steps_interval[interval_max, ]$step)

# Averages are calculated at the 5 minute interval for all of the days here. 
```
## Imputing missing values
```{r}
null_steps <- sum(is.na(as.character(activity$steps)))
null_interval <- sum(is.na(as.character(activity$interval)))
null_date <- sum(is.na(as.character(activity$date)))
null_total <- sum(!complete.cases(activity))
StepsAverage <- aggregate(steps ~ interval, activity, mean)
fill_nulls <- numeric()

for (i in 1:nrow(activity)) {
  obs <- activity[i, ]
  if (is.na(obs$steps)) {
    steps <- subset(StepsAverage, interval == obs$interval)$steps
  } else {
    steps <- obs$steps
  }
  fill_nulls <- c(fill_nulls, steps)
}
data_clone <- activity
data_clone$steps <- fill_nulls
```
# Uses mean imputation for NA values to have filled in. 

## Are there differences in activity patterns between weekdays and weekends?
```{r}
steps_day2 <- aggregate(steps ~ date, data_clone, sum)
hist(steps_day2$steps, main = paste("Total Steps Taken Each Day"), col="red", xlab="Number of Steps")
hist(steps_day$steps, main = paste("Total of Steps Taken Each Day"), col="blue", xlab="Number of Steps", add=T)
legend("topleft", c("New dataset", "Dataset"), col=c("red", "blue"), lwd=10)
mean(steps_day2$steps)
median(steps_day2$steps)
median1 <- median(steps_day$steps)
median2 <- median(steps_day2$steps)
median2 - median1
wkday <- function(dat_val) {
  wd <- weekdays(as.Date(dat_val, '%Y-%m-%d'))
  if  (!(wd == 'Saturday' || wd == 'Sunday')) {
    x <- 'Weekday'
  } 
  else {
    x <- 'Weekend'
  }
  x
}
activity$Day <- as.factor(sapply(activity$date, wkday))
report_activity <- aggregate(steps~interval+Day,activity,mean)
g<- ggplot(report_activity, aes(interval,steps))
g <-g+ geom_line(stat = 'identity') + facet_grid(Day~.)
g+ labs(x= '5 mins Interval', y = "Average of Steps") + ggtitle("Weekday versus Weekend")
```
# Yes, there appear to be differences between the weekday and the weekend. Weekend activities appear to be more consistent, where the weekday has more volatility and greater variance. 
